Getting Started with the Graded Response Model (GRM): A gentle introduction and tutorial in R

### 2023-11-23 draft_paper.qmd
Total overhaul of the R script. I used a wrong formula to reverse unfav items.

---
title: "Getting Started with the Graded Response Model (GRM): A gentle introduction and tutorial in R"
shorttitle: "Graded Response Model Tutorial"
author:
  - name: Rizqy Amelia Zein
    corresponding: true
    orcid: 0000-0001-7840-0299
    email: Amelia.Zein@psy.lmu.de
    affiliations:
      - name: Ludwig-Maximilians-Universität
        department: Department of Psychology
        address: Leopoldstraße 13
        city: Munich
        postal-code: "80802"
      - name: Universitas Airlangga
        department: Department of Psychology
        address: Jalan Airlangga 4-6
        city: Surabaya
        region: Jawa Timur, Indonesia
        postal-code: "60286"
  - name: Hanif Akhtar
    orcid: 0000-0002-1388-7347
    affiliations:
      - name: ELTE Eötvös Loránd University
        department: Doctoral School of Psychology
        country: Hungary
      - name: Universitas Muhammadiyah Malang
        department: Faculty of Psychology
        city: Malang
        region: Jawa Timur, Indonesia
abstract: "This tutorial paper introduces the Graded Response Model (GRM), a tool for testing measurement precision under the Item Response Theory (IRT) paradigm. Addressing common problems of measurement imprecision and lack of construct validity, the tutorial guides researchers through a one-dimensional GRM analysis in the R environment, using psych, mirt, and ggmirt packages. GRM is specifically designed to examine the psychometric properties of psychological scales with polytomous items. The tutorial illustrates the procedure using data from the Open Psychometrics Database on the Right-Wing Authoritarianism (RWA) scale, outlining the theoretical underpinnings of GRM, and steps for data preparation, model fitting, interpretation of results, and dealing with common issues and anomalies that may typically arise in the process."
keywords:
  - graded response model
  - item response theory
  - mirt
  - ggmirt
  - psych
  - R
author-note:
  disclosures:
    conflict-of-interest: "We declare we have no competing interests."
    financial-support: "The first author receives a PhD scholarship from the German Academic Exchange Service (Deutscher Akademischer Austauschdienst, DAAD)."
lang: en
bibliography: references.bib
format:
  apaquarto-docx: default
editor: visual
---

@@

```{r, echo=F, include=F}
#| label: setup
# install.packages("tidyverse", "psych", "devtools", "mirt", "caret", dependencies=TRUE) 
# INFO: Remove the hashtag before install.package() to start installing packages (only if you have not had them installed)
# devtools::install_github("masurp/ggmirt")
# Now, we activate the package:
library(tidyverse); library(psych); library(mirt); library(ggmirt); library(caret)

@@

@@

>>>>

---
title: "Getting Started with the Graded Response Model (GRM): A gentle introduction and tutorial in R"
shorttitle: "A Graded Response Model Tutorial"
author:
  - name: Rizqy Amelia Zein
    corresponding: true
    orcid: 0000-0001-7840-0299
    email: amelia.zein@psikologi.unair.ac.id
    affiliations:
      - name: Universitas Airlangga
        department: Department of Psychology
        address: Jalan Airlangga 4-6
        city: Surabaya
        region: Jawa Timur, Indonesia
        postal-code: "60286"
      - name: Ludwig-Maximilians-Universität
        department: Department of Psychology
        address: Leopoldstraße 13
        city: Munich
        postal-code: "80802"
  - name: Hanif Akhtar
    orcid: 0000-0002-1388-7347
    affiliations:
      - name: ELTE Eötvös Loránd University
        department: Doctoral School of Psychology
        country: Hungary
      - name: Universitas Muhammadiyah Malang
        department: Faculty of Psychology
        city: Malang
        region: Jawa Timur, Indonesia
abstract: "This tutorial paper introduces the Graded Response Model (GRM), a tool for testing measurement precision under the Item Response Theory (IRT) paradigm. Addressing common problems of measurement imprecision and lack of construct validity, the tutorial guides researchers through a one-dimensional GRM analysis in the R environment, using psych, mirt, and ggmirt packages. GRM is specifically designed to examine the psychometric properties of psychological scales with polytomous items. The tutorial illustrates the procedure using data from the Open Psychometrics Database on the Right-Wing Authoritarianism (RWA) scale, outlining the theoretical underpinnings of GRM, and steps for data preparation, model fitting, interpretation of results, and dealing with common issues and anomalies that may typically arise in the process."
keywords:
  - graded response model
  - item response theory
  - mirt
  - ggmirt
  - psych
  - R
author-note:
  disclosures:
    conflict-of-interest: "We declare we have no competing interests."
    financial-support: "The first author receives a PhD scholarship from the German Academic Exchange Service (Deutscher Akademischer Austauschdienst, DAAD)."
lang: en
bibliography: references.bib
format:
  apaquarto-docx: default
editor: visual
execute: 
  warning: false
  error: false
  output: false
  echo: true
---

@@

```{r}
#| include: false
#| label: "setup"

# install.packages("flextable", dependencies=TRUE) 
# INFO: Remove the hashtag before install.package() to start installing packages (only if you have not had them installed)
library(flextable)

@@

Measurement, undoubtedly, is an important foundation of scientific research. Measurement process start with defining the construct being studied and then determine the assumptions

# A Brief Overview of Item Response Theory

# Graded Response Model

# Assumptions and Key Concepts

# Disclosure

To maximise reproducibility of our analysis, we wrote the article as a Quarto (.qmd) document, where we integrate the R codes used in the analysis as well as its outputs. We also include the complete R script for the example we used as a supplementary document. The Quarto file (and its corresponding .docx and .pdf output) and R script are publicly available on (a Github repository)\[https://github.com/rameliaz/grm-tutorial-paper\].

# An Illustrative Example of Graded Response Model: The Right Wing Authoritarianism (RWA) Scale

To demonstrate the procedure of running a graded response model, we

## A Brief Overview of the Altemeyer's RWA Scale

## Step 1: Preparation

```{r}

install.packages("tidyverse", "psych", "devtools", "mirt", "caret", dependencies=TRUE)
library(tidyverse); library(psych); library(mirt); library(caret)
```

Blabla

```{r}

# devtools::install_github("masurp/ggmirt")
library(ggmirt)
```

Blabla

```{r}
ds <- read.csv("data/data.csv")
```

Blabla

```{r}
str(ds)
```

Blabla

```{r}
rwa <- subset(ds, select = Q1:Q22)
str(rwa)
```

Blabla

## Step 2: Inspecting Key Descriptive Statistics

```{r}
psych::describe(rwa)
```

As we see in {apatb-table1}, nddasjdhajskdhaks

```{r}
unfav <- c("Q4","Q6","Q8","Q9","Q11","Q13","Q15","Q18","Q20","Q21") # Now we create a vector defining which items will be coded reversely.
rwa <- rwa %>% 
  mutate(across(all_of(unfav), ~ 9 - .))# We simply subtract the scores from 9 (the maximum) to reverse code the unfavorable items.
```

Kok ada nilai 0-nya? padahal kan skor minimalnya 1. Coba kita hitung frekuensi nilai 0 di tiap item.

```{r}
zero <- colSums(rwa == 0) / nrow(rwa) * 100 # Computing the frequency of "0" in each column.
print(zero) # The proportion of "0" for each item.
```

```{r}
rwa <- rwa %>%
  mutate_all(~na_if(., 0)) %>%  # Replacing 0 with NA in all columns.
  drop_na()  # Removing cases with any NA values.
```

Now let's look again at the descriptive statistics after we removed all NA cases.

```{r}
psych::describe(rwa)
```

## Step 3: Examining Dimensionality

```{r}
#| eval: false

irt.fa(rwa, nfactors = 1, fm = "minres")
```

```{r}
#| eval: false

fa.parallel(rwa, nfactors = 1, fm="minres", fa="fa", cor = "poly")
```

```{r}
cor <- cor(rwa,method="pearson") # First, creating a (pearson) correlation matrix.
efa <- fa(rwa, nfactors=1, fm="minres") # Now, running exploratory factor analysis. 
print(efa) # Print the results.
```

Scree plot

Now lets do parallel analysis

```{r}
pa <- fa.parallel(rwa, fm="minres", fa="fa") # Running a parallel analysis.
pa$fa.values # Seeing the eigenvalues of each factor.
```

## Step 4: Model Estimation, Parameters, and Fit Statistics

```{r}
model <- 'rwa = 1-22'
```

```{r}
fit <- mirt(data=rwa, 1, model=model, itemtype="graded", SE=T, verbose=F)
```

```{r}
coefs <- coef(fit, IRTpars=T, simplify=T) # Storing model parameters in a data frame.
print(coefs) # Yielding model parameters: item discriminations (a) and threshold (b).
```

```{r}
summary(fit)
```

lhalalala

```{r}
M2(fit, type="C2")
```

lhalalala

```{r}
#| include: false

m2_results <- M2(fit, type="C2")
m2_stat <- m2_results$M2
df <- m2_results$df
p_value <- m2_results$p.value
rmsea <- m2_results$RMSEA
srmr <- m2_results$SRMR
tli <- m2_results$TLI
cfi <- m2_results$CFI
```

According to our analysis, the model does not fit the data well (*M*(`r df`) = `r m2_stat`, *p* = `r p_value`, *RMSEA* = `r rmsea`, )

```{r}
item.fit <- itemfit(fit)
```

## Step 5: Model Residuals

```{r}
ld <- residuals(fit, type = "LD") # Running local dependency statistics
up <- which(upper.tri(ld), arr.ind = T) # Extracting values only on the upper side of the diagonal.
lar <- up[ld[up] > 0.2 | ld[up] < -0.2, ] # Defining unusually large residuals (>0.2).
```

lhalalala

```{r}
for (i in 1:nrow(lar)) {
  row <- lar[i, 1]
  col <- lar[i, 2]
  value <- ld[row, col]
  cat(sprintf("A large residual correlation is found between item %d and item %d: %f\n", row, col, value))
} # Now we detect the problematic pairs.
```

lhalalala

```{r}
q3 <- residuals(fit, type = "Q3") # Running Yen's Q3 statistics
findCorrelation(q3, cutoff = 0.2, verbose = T) # Detecting problematic correlation pairs.
```

## Step 6: IRT Plots

Now spit out the

```{r}
tracePlot(fit, facet=T, title = "Category Probability Functions of RWA Scale") + labs(color="Response Options")
```

lhalalala

```{r}
itemInfoPlot(fit, facet=T, title = "Item Information Curves of the RWA Scale")
```

lhalalala

```{r}
#| output: false

itemInfoPlot(fit, facet=T, theta_range = c(-6,6), title = "Item Information Curves of the RWA Scale")#
```

lhalalala

```{r}
testInfoPlot(fit, title="Test Information Curve of the RWA Scale")
```

## Step 7: Computing Reliability

```{r}
theta_se <- fscores(fit, full.scores.SE = T) # Extracting the estimated theta score of each participant.
e_rel <- empirical_rxx(theta_se) # Then use the estimated theta to calculate empirical reliability.
```

lhalalala

```{r}
m_rel <- marginal_rxx(fit)
```

lhalalala

```{r}
omega(rwa)
```

# Conclusions

@@

```{r apatb-table1}
#| output: true
#| echo: false
#| apa-cap: Descriptive Statistics of RWA Scale
#| apa-note: SD = Standard Deviation
#| ft.align: left

# Transforming describe() output into a APA-formatted table
table1 <- psych::describe(rwa); table1 <- as.data.frame(table1)
table1 <- table1 %>% 
  rownames_to_column(var="Item") %>% 
  select(Item, mean, sd, min, max, range, n) %>% 
  rename(Mean=mean, SD=sd, Minimum=min, Maximum=max, Range=range) %>% 
  flextable() %>%
  theme_apa() %>% 
  line_spacing(part = "all") %>% 
  padding(padding.top = 5, padding.bottom = 5)

table1 <- colformat_double(x=table1, j="Minimum", digits = 0)
table1 <- colformat_double(x=table1, j="Maximum", digits = 0)
table1 <- colformat_double(x=table1, j="Range", digits = 0)
table1 <- colformat_double(x=table1, j="n", digits = 0)

table1
```

```{r apatb-table2}
#| output: true
#| echo: false
#| apa-cap: Descriptive Statistics of RWA Scale
#| apa-note: SD = Standard Deviation, Descriptive Statistics After Reversing Unfavorable Items
#| ft.align: left

# Transforming describe() output into a APA-formatted table
unfav <- c("Q4","Q6","Q8","Q9","Q11","Q13","Q15","Q18","Q20","Q21") # Now we create a vector defining which items will be coded reversely.
rwa <- rwa %>% 
  mutate(across(all_of(unfav), ~ 9 - .))# We simply subtract the scores from 9 (the maximum) to reverse code the unfavorable items.

table2 <- psych::describe(rwa); table2 <- as.data.frame(table2)
table2 <- table2 %>% 
  rownames_to_column(var="Item") %>% 
  select(Item, mean, sd, min, max, range, n) %>% 
  rename(Mean=mean, SD=sd, Minimum=min, Maximum=max, Range=range) %>% 
  flextable() %>% 
  theme_apa() %>% 
  line_spacing(part = "all") %>% 
  padding(padding.top = 5, padding.bottom = 5)

table2 <- colformat_double(x=table2, j="Minimum", digits = 0)
table2 <- colformat_double(x=table2, j="Maximum", digits = 0)
table2 <- colformat_double(x=table2, j="Range", digits = 0)
table2 <- colformat_double(x=table2, j="n", digits = 0)

table2
```

```{r apatb-table3}
#| output: true
#| echo: false
#| apa-cap: Percentge of Zero in Each Item
#| ft.align: left

table3 <- as.data.frame(zero)

table3 <- table3 %>% 
  rownames_to_column(var="Item") %>% 
  select(Item, zero) %>% 
  rename(Percentage=zero) %>% 
  flextable() %>% 
  theme_apa() %>% 
  line_spacing(part = "all") %>% 
  padding(padding.top = 5, padding.bottom = 5)

table3

```{r apatb-table4}
#| output: true
#| echo: false
#| apa-cap: Descriptive Statistics of RWA Scale
#| apa-note: SD = Standard Deviation, Descriptive Statistics After Reversing Unfavorable Items
#| ft.align: left

rwa <- rwa %>%
  mutate_all(~na_if(., 0)) %>%  # Replacing 0 with NA in all columns.
  drop_na()  # Removing cases with any NA values.

# Transforming describe() output into a APA-formatted table
table4 <- psych::describe(rwa); table4 <- as.data.frame(table4)
table4 <- table4 %>% 
  rownames_to_column(var="Item") %>% 
  select(Item, mean, sd, min, max, range, n) %>% 
  rename(Mean=mean, SD=sd, Minimum=min, Maximum=max, Range=range) %>% 
  flextable() %>% 
  theme_apa() %>% 
  line_spacing(part = "all") %>% 
  padding(padding.top = 5, padding.bottom = 5)

table4 <- colformat_double(x=table4, j="Minimum", digits = 0)
table4 <- colformat_double(x=table4, j="Maximum", digits = 0)
table4 <- colformat_double(x=table4, j="Range", digits = 0)
table4 <- colformat_double(x=table4, j="n", digits = 0)

table4
```

```{r fig1}
#| output: true
#| echo: false
#| label: apafg-fig1
#| apa-cap: Scree Plot
#| apa-note: NULL

plot(efa$values, type = "b", main = "Scree Plot", xlab = "Factor", ylab = "Eigenvalue") # Scree plot
abline(h = 1, col = "red", lty = 2) # Add new line to factor 1.

```{r fig2}
#| output: true
#| echo: false
#| label: apafg-fig2
#| apa-cap: Parallel Analysis
#| apa-note: NULL

plot(pa)
```

```{r apatb-table5}
#| output: true
#| echo: false
#| apa-cap:  Descriptive Statistics of RWA Scale
#| apa-note: SD = Standard Deviation, Descriptive Statistics After Reversing Unfavorable Items
#| ft.align: left

table5 <- as.data.frame(coefs$items)

table5 <- table5 %>% 
  rownames_to_column(var="Item") %>% 
  select(Item, a, b1, b2, b3, b4, b5, b6, b7, b8) %>% 
  rename(α=a, β1=b1, β2=b2, β3=b3, β4=b4, β5=b5, β6=b6, β7=b7, β8=b8) %>% 
  flextable() %>% 
  theme_apa() %>% 
  line_spacing(part = "all") %>% 
  padding(padding.top = 5, padding.bottom = 5)

table5 <- colformat_double(x=table5, j="α", digits = 2)
table5 <- colformat_double(x=table5, j="β1", digits = 2)
table5 <- colformat_double(x=table5, j="β2", digits = 2)
table5 <- colformat_double(x=table5, j="β3", digits = 2)
table5 <- colformat_double(x=table5, j="β4", digits = 2)
table5 <- colformat_double(x=table5, j="β5", digits = 2)
table5 <- colformat_double(x=table5, j="β6", digits = 2)
table5 <- colformat_double(x=table5, j="β7", digits = 2)
table5 <- colformat_double(x=table5, j="β8", digits = 2)

table5

```{r apatb-table6}
#| output: true
#| echo: false
#| apa-cap: Descriptive Statistics of RWA Scale
#| apa-note: SD = Standard Deviation, Descriptive Statistics After Reversing Unfavorable Items
#| ft.align: left

table6 <- as.data.frame(item.fit)

table6 <- table6 %>%
  select(item, S_X2, df.S_X2, RMSEA.S_X2, p.S_X2) %>% 
  rename(Item=item) %>% 
  flextable() %>% 
  theme_apa() %>% 
  line_spacing(part = "all") %>% 
  padding(padding.top = 5, padding.bottom = 5)

table6 <- colformat_double(x=table6, j="S_X2", digits = 2)
table6 <- colformat_double(x=table6, j="df.S_X2", digits = 2)
table6 <- colformat_double(x=table6, j="RMSEA.S_X2", digits = 2)
table6 <- colformat_double(x=table6, j="p.S_X2", digits = 2)

table6

```{r fig3}
#| output: true
#| echo: false
#| label: apafg-fig3
#| apa-cap: Category Probability Functions of RWA Scale
#| apa-note: NULL

tracePlot(fit, facet=T, title = "Category Probability Functions of RWA Scale") + labs(color="Response Options")
```

```{r fig4}
#| output: true
#| echo: false
#| label: apafg-fig4
#| apa-cap: Item Information Curves of the RWA Scale
#| apa-note: NULL

itemInfoPlot(fit, facet=T, title = "Item Information Curves of the RWA Scale")
```

```{r fig5}
#| output: true
#| echo: false
#| label: apafg-fig5
#| apa-cap:  Item Information Curves of the RWA Scale Ranging from -6 to +6 Standard Deviation
#| apa-note: NULL

itemInfoPlot(fit, facet=T, theta_range = c(-6,6), title = "Item Information Curves of the RWA Scale")#
```

```{r fig6}
#| output: true
#| echo: false
#| label: apafg-fig6
#| apa-cap: Test Information Curve of the RWA Scale
#| apa-note: NULL

testInfoPlot(fit, title="Test Information Curve of the RWA Scale")
```

```{r fig7}
#| output: true
#| echo: false
#| label: apafg-fig7
#| apa-cap: Reliability of the RWA Scale Given to the θ Level
#| apa-note: NULL

conRelPlot(fit, title="Reliability of the RWA Scale Given to the θ Level")
```
